\chapter{Acknowledgements}

I am very grateful to countless people who have assisted with and contributed to this dissertation in their own way.

Firstly, I want to thank my advisor, Prof. Xianzhe Jia for giving me this opportunity to pursue scientific research as a career. His advice, professionalism, support and selfless assistance have been invaluable for the past six years. I could not have asked for a better role-model and mentor as an academic advisor. I have learned a lot working with him and have had the most positive experience as part of his research group. 

I am very grateful to Prof. James Slavin for his mentorship, and for being so generous with his time. His encouragement and enthusiasm have made this journey much easier, and I have learned a lot from our interactions. 

This work would not have been possible without the support of Prof. Gabor Toth, whose technical guidance and experience with the SWMF code has been crucial. I am very grateful for all his assistance and prompt response to any questions or issues.

I would like to thank Prof. Mark Kushner, who took time out of his busy schedule to serve as the cognate member of the committee. 

This thesis and my graduate experience has greatly benefited from conversations and assistance from many colleagues. In particular I would like to thank Camilla Harris, Ryan Dewey, Sergio Vidal-Luengo, Abigail Azari, Jamie Jasinski and many other graduate students and postdocs for their kindness and interesting conversations. I'm also grateful to Yuxi Chen, who was always prompt and comprehensive in his help, whenever I had any issues with the SWMF.  

I am indebted to the CLASP IT group - Faye, Darren, Bryan, and Cameron. During the initial months of the pandemic, while the rest of us were sequestered at home, they would manually restart my office computer every few days while it suffered from unexplained crashes. They also tried their utmost to recover data which was lost due to a malfunctioning hard drive. I am truly grateful for their support. I would also like to thank Laura Hopkins, Claire Miller and Sandra Pytlinski for making every administration related task seem smooth. 

I am grateful for the support of my friends from India and the US. In particular, I would like to thank Ayoub and Sergio for their friendship, frank yet constructive criticism, motivation and for many interesting discussions. 

My experience in graduate school has been quite painless due to the constant support, honest feedback and patience of my partner, Bishakha. I am extremely grateful to my grandmothers, to whom this thesis is dedicated, for their support and commitment to learning and knowledge. Most of all, I thank my brother and my parents, who have been a constant source of support and strength, and who gave me the precious freedom and encouragement to pursue my interests; a risk not many people can afford to take.